---
title: "Class Reference"
author: "Kriti Singh"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).

```{r}
setwd("~/jour472/data_journalism_2024_fall")
```
Here's how to add code: 
```{r}
2+2
```

### How to install and load the tidyverse

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### Adding readr and dpylr
```{r}
library(readr)
library(dplyr)
library(tidyverse)
```

### importing data
```{r}
maryland_winred_contributions <- read_rds("https://thescoop.org/files/maryland_winred.rds")
```

### glimpse: function to get an overview of the data, the metadata of the data (remember to run tidyverse every single time)
```{r}
glimpse(maryland_winred_contributions)
```
### head: returns first 6 rows by default, can return n if head(x,n=number), x = input, dataset n = number of rows head should return
```{r}
head(maryland_winred_contributions)
```
### counting number of contributions to each recipient

### First it takes all contributions: maryland_winred_contributions, next, groups are created for each unique contributions for the same candidate, summary then performs a summary operation for each group created, ie., here, counts number of rows in each group

### difference between group_by and summarise: splits the data into groups on one or more variables. does not change the data itself, changes how subsequent operations are applied to the data. summarize, (i think) works on the groups created by group_by
```{r}
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n()
  )
```
### group by “memo_text” and “date” to see how many contributions occurred on every date to every recipient. sorting by the count of contributions in descending order.
```{r}
maryland_winred_contributions |>
  group_by(memo_text, date) |>
  summarise(
    count_contribs = n()
  ) |>
  arrange(desc(count_contribs))
```
### summarizing using mean, median
```{r}
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) |>
  arrange(desc(max_amount))
```
### sanity check
click on the broom icon and remove all objects. 

### things to remember
read_csv is a tidyverse function, understands what data types are there. better version of read.csv
r sorts data not just alphabetically but also acc to case 
read.csv is a base R function (forget this exists)
clean the dataset and turn it into a shape you like and THEN work with it