---
title: "Data Analysis Project"
author: "Group 4"
date: "11-16-2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Deliverable 2 (week 12) - 10 percent*: 

*Question 1* Lists your questions - if i asked you to refine them, show me that you've done that.

*Answer 1* 
Q1. For the big picture, in *Maryland*,
1.1 How many voters registered for each party in the 2024 primary and general election? 
1.1.1 How have these numbers changed compared to those in previous elections?
1.2 Compared to the total population of the state, what is the proportion of voters who register for the election?
1.3 Compared to the 2024 presidential election result, does the registration pattern match with the result? 
---
Q2. At the *Congressional District level*
2.1 How many voters registered for each party in the 2024 primary and general election?
2.1.1 How have these numbers changed compared to those in previous elections?
2.2 Are there any patterns for voter registration from every district?
2.2.1 Are there any districts that don't fall into this pattern? 
2.2.2 How are those districts different from others in terms of location, demography, economy, etc.?
2.3 Compared to the total population of each district, what is the proportion of voters who register for the election? 
2.3.1 What are the differences between precincts having large and small proportions of voter registration?
2.4 Compared to the representative race result, does the registration pattern match with the result? 
---
Q3. At the *Legislative Districts level*
3.1 How many voters registered for each party in the 2024 primary and general election?
3.1.1 How have these numbers changed compared to those in previous elections?
3.2 Are there any patterns for voter registration from every district?
3.2.1 Are there any districts that don't fall into this pattern? 
3.2.2 How are those districts different from others in terms of location, demography, economy, etc.?
3.3 Compared to the total population of each district, what is the proportion of voters who register for the election? 
3.3.1 What are the differences between precincts having large and small proportions of voter registration?
3.4 Compared to the election result, does the registration pattern match with the result? 
---
Q4. At the *Precinct level*
4.1 How many voters registered for each party in the 2024 primary and general election?
4.1.1 How have these numbers changed compared to those in previous elections?
4.2 Are there any patterns for voter registration from every precinct?
4.2.1 Are there any precincts that don't fall into this pattern? 
4.2.2 How are those precincts different from others in terms of location, demography, economy, etc.?
4.3 Compared to the total population of each precinct, what is the proportion of voters who register for the election? 
4.3.1 What are the differences between precincts having large and small proportions of voter registration? 
4.4 Compared to the election result, does the registration pattern match with the result?
---
Q5. At the *County level*
5.1 How many voters registered for each party in the 2024 primary and general election?
5.1.1 How have these numbers changed compared to those in previous elections?
5.2 Are there any patterns for voter registration from every county?
5.2.1 Are there any counties that don't fall into this pattern? 
5.2.2 How are those counties different from others in terms of location, demography, economy, etc.?
5.3 Compared to the total population of each county, what is the proportion of voters who register for the election? 
5.3.1 What are the differences between counties having large and small proportions of voter registration? 
5.4 Compared to the 2024 presidential election result, does the registration pattern match with the result? 

*Question 2* Loads and cleans the core data set to prepare for analysis.

*Answer 2*

```{r}
#Load Libraries 
library(tidyverse)
library(janitor)
library(tidycensus)
library(rvest)
library(dplyr)
```


```{r}
#Prep dataframe for Q1 focusing on Maryland

#Registration Data
md_primary_24 <- read_csv("data/county/county_primary_24.csv")
md_general_24 <- read_csv("data/county/county_general_24.csv")
md_primary_20 <- read_csv("data/county/county_primary_20.csv")
md_primary_20 <- md_primary_20 |>
  rename(County = 1)
md_general_20 <- read_csv("data/county/county_general_20.csv")

#Population Data
state_population <- get_acs(geography="state", 
                       variables = "B01003_001", 
                       year=2022)
md_population <- state_population |>
  slice(21)

#Presidential Election Result
pres_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_1.html"
pres_election_result_html <- pres_election_result_url |>
  read_html() |>
  html_table()
pres_election_result <- pres_election_result_html[[1]]

```
```{r}
#Prep dataframe for Q2 focusing on congressional district level

#Registration Data
congress_primary_24 <- read_csv("data/congress_level/congress_primary_24.csv")
congress_general_24 <- read_csv("data/congress_level/congress_general_24.csv")
congress_primary_20 <- read_csv("data/congress_level/congress_primary_20.csv")
congress_primary_20 <- congress_primary_20 |>
  rename(District = 1) |>
  rename(county = 2) |>
  rename(BAR = 3) |>
  rename(DEM = 4) |>
  rename(OTH = 5) |>
  rename(REP = 6) |>
  rename(UNA = 7) |>
  rename(WCP = 8) |>
  rename(total = 9) 
congress_general_20 <- read_csv("data/congress_level/congress_general_20.csv")

#Population Data
congress_population <- get_acs(geography="congressional district", 
                       variables = "B01003_001", state = "MD", 
                       year=2022)

#Representative Election Result
representative_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_3.html"
representative_election_result_html <- representative_election_result_url |>
  read_html() |>
  html_table()
representative_election_result1 <- representative_election_result_html[[1]]
representative_election_result2 <- representative_election_result_html[[2]]
representative_election_result3 <- representative_election_result_html[[3]]
representative_election_result4 <- representative_election_result_html[[4]]
representative_election_result5 <- representative_election_result_html[[5]]
representative_election_result6 <- representative_election_result_html[[6]]
representative_election_result7 <- representative_election_result_html[[7]]
representative_election_result8 <- representative_election_result_html[[8]]
```
```{r}
#Prep dataframe for Q3 focusing on legislative district level

#Registration Data
legis_primary_24 <- read_csv("data/legis_level/legis_primary_24.csv")
legis_general_24 <- read_csv("data/legis_level/legis_general_24.csv")
legis_primary_20 <- read_csv("data/legis_level/legis_primary_20.csv")
legis_general_20 <- read_csv("data/legis_level/legis_general_20.csv")

#Population Data
legis_population <- get_acs(geography="state legislative district (lower chamber)", 
                       variables = "B01003_001", state = "MD", 
                       year=2022)

#Presidential Election Result 
#Download Precinct Result and Code
precint_result <- read_csv("data/legis_level/result/result.csv")
legis_code <- read_csv("data/legis_level/result/code.csv")
#Prep result dataframe
precint_result <- precint_result |>
  filter(Contest == "President and Vice President of the United States")
precint_result <- precint_result |>
  select(2,4,6,7) |>
  clean_names() |>
  mutate(precinct_name = as.numeric(gsub("-","", precinct_name)))
#Prep code dataframe
legis_code <- legis_code |>
  select(3,4,10) |>
  filter(!is.na(ED_PRECINCT)) |>
  filter(!is.na(LEGISLATIVE_SUB_DISTRICT_CODE))
#Join result dataframe with code dataframe to generate legislative district results 
legis_result_draft <- precint_result |> 
  left_join(legis_code, join_by(jurisdiction==COUNTY_NAME, precinct_name==ED_PRECINCT)) |>
  filter(!is.na(LEGISLATIVE_SUB_DISTRICT_CODE)) |>
  group_by(LEGISLATIVE_SUB_DISTRICT_CODE, party) |>
  summarise(
    total_votes = sum(total_votes))
#Reshape dataframe
legis_result_DEM <- legis_result_draft |>
  filter(party == "DEM") |>
  select(1,3) |>
  rename(DEM_result = 2)
legis_result_GRN <- legis_result_draft |>
  filter(party == "GRN") |>
  select(3) |>
  rename(GRN_result = 2) 
legis_result_LIB <- legis_result_draft |>
  filter(party == "LIB") |>
  select(3) |>
  rename(LIB_result = 2)
legis_result_NON <- legis_result_draft |>
  filter(party == "NON") |>
  select(3) |>
  rename(NON_result = 2)
legis_result_REP <- legis_result_draft |>
  filter(party == "REP") |>
  select(3) |>
  rename(REP_result = 2)
legis_result_UNF <- legis_result_draft |>
  filter(party == "UNF") |>
  select(3) |>
  rename(UNF_result = 2)


legis_result <- bind_cols(legis_result_DEM, legis_result_GRN, legis_result_LIB, legis_result_NON, legis_result_REP, legis_result_UNF) |>
  select(-3,-5,-7,-9,-11) |>
  rename(legislative_district_code = 1)

```

```{r}
#Prep dataframe for Q4 focusing on legislative precinct level

#I have that for this election from the state, and should have it for 2022 (will need to check). Start with county and district stuff first and we'll figure out if the precinct-level analysis will work.
```


```{r}
#Prep dataframe for Q5 focusing on county level

#Registration Data
county_primary_24 <- read_csv("data/county/county_primary_24.csv")
county_general_24 <- read_csv("data/county/county_general_24.csv")
county_primary_20 <- read_csv("data/county/county_primary_20.csv")
county_primary_20 <- county_primary_20 |>
  rename(County = 1)
county_general_20 <- read_csv("data/county/county_general_20.csv")

#Population Data
county_population <- get_acs(geography="county", 
                       variables = "B01003_001", 
                       state = "MD", year=2019)

#Presidential Election Result (county level)
pres_election_result_county_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_detail_results_2024_1_1.html"
pres_election_result_county_html <- pres_election_result_county_url |>
  read_html() |>
  html_table()
pres_election_result_county1 <- pres_election_result_county_html[[1]]
pres_election_result_county2 <- pres_election_result_county_html[[2]]
pres_election_result_county2 <- pres_election_result_county2|>
  select(-1)
pres_election_result_county3 <- pres_election_result_county_html[[3]]
pres_election_result_county3 <- pres_election_result_county3|>
  select(-1)
pres_election_result_county4 <- pres_election_result_county_html[[4]]
pres_election_result_county4 <- pres_election_result_county4|>
  select(-1)
pres_election_result_county5 <- pres_election_result_county_html[[5]]
pres_election_result_county5 <- pres_election_result_county5|>
  select(-1)
pres_election_result_county6 <- pres_election_result_county_html[[6]]
pres_election_result_county6 <- pres_election_result_county6|>
  select(-1)
pres_election_result_county <- bind_cols(pres_election_result_county1, pres_election_result_county2, pres_election_result_county3, pres_election_result_county4, pres_election_result_county5, pres_election_result_county6)
```
*Question 3* Shows basic exploratory analysis to demonstrate an understanding of the dataset, including the number of rows and columns, any obvious limitations or flaws and any reasons why it might not be able to answer the questions you've posed last week.

*Answer 3*
Q1
We may not be able to compare the change in the number of registrations for LIB and NLM parties (from primary), and NLM parties (from general) between 2024 and 2020 as the dataframes from 2020 don’t have data for these parties.
Q2
We may not be able to compare the change in the number of registrations at the congressional level for LIB and NLM parties (from primary), and NLM party (from general) between 2024 and 2020 as the dataframes from 2020 don’t have data for these parties. 
Q3
We may not be able to compare the change in the number of registrations at the legislative level for the NLM party (from both primary and general) between 2024 and 2020 as the dataframes from 2020 don’t have data for this party. 
Q4
Q5
- We may not be able to compare the change in the number of registrations for LIB and NLM parties (from primary), and NLM parties (from general) between 2024 and 2020 as the dataframes from 2020 don’t have data for these parties.

*Question 4* Describe the limits of the data in terms of timeframe, what is and is not included and any codes or other information you'd need to find definitions for, and what you've done to better define it.

*Answer 4*
Q1
- The registration breakdown by county doesn’t have the total row of registrations for each party which we have to use to answer questions at the state level. We may need to find the sum ourselves.
- The population data of Maryland is not included here. We have to download the data from the census.
- We are not sure how to make the comparison between the voter registration and the election result data (https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_1.html) makes sense because the registration is categorized by parties but the result is categorized by candidates which some of them come from the same party.    
Q2
- We’re not sure whether there is population data for each congressional district, so the part of this question about comparing registration with population may not be able to answer. 
Q3
- We’re not sure whether there is population data for each legislative district, so the part of this question about comparing registration with population may not be able to answer. 
- We’re not sure which election result we should use to compare with the registration data at the legislative level. 
Q4
- We’re not sure where or how to locate registration at the precinct level.
Q5
- The population data of each county in Maryland is not included here. We have to download the data from the census.
- We are not sure how to make the comparison between the voter registration and the election result data (https://elections.maryland.gov/elections/2024/general_Results/gen_detail_results_2024_1_1.html) makes sense because the registration is categorized by parties but the result is categorized by candidates which some of them come from the same party.

-------------------------------------------------------------------------------------------------

*Deliverable 3 (week 13):* 

Turn in, via GitHub, a draft of the data notebook you will produce for this project. You should work on this notebook together, but everyone needs to upload a copy to their own GitHub repo and turn in a link for grading. At a minimum, your data notebook should load the data you'll use with this project and have demonstrated work on at least two of the questions you are seeking to answer. The more you have finished, the more feedback I'll be able to give you.

Q1. For the big picture, in *Maryland*,
1.1 How many voters registered for each party in the 2024 primary and general election?
*Answer: In the 2024 general elections, the registration totals were as follows: 2211316 registered for the Democratic Party, 1007384 for the Republican Party, 78234 for other parties, and 907638 were unaffiliated. In the primary elections, 2202830 registered for the Democrats, 994388 for the Republicans, 31418 for other parties, and 451465 were unaffiliated.*
1.1.1 How have these numbers changed compared to those in previous elections?
*Answer: Compared to the previous election (2020, the primary elections in 2024 saw a decline in registrations for the major parties, while other parties and unaffiliated registrants gained significant popularity. Specifically, the Democratic Party lost 1.24%, the Republican Party lost 1.45%, while other parties gained 44.95%, and unaffiliated registrations increased by 40.24%. This trend continued in the general elections, with the Democratic Party experiencing a 2.27% decrease, while the Republican Party saw a smaller drop of 0.82%. Similarly, other parties gained 21.54%, and unaffiliated registrations rose by 18.36%.*
1.2 Compared to the total population of the state, what is the proportion of voters who register for the election?
*Answer: In the primary elections, voter turnout was 59.72% of the total population, while in the general elections, it increased to 68.23%.*
1.3 Compared to the 2024 presidential election result, does the registration pattern match with the result?
*Answer: By comparing the general party registrations in Maryland for 2024 with the presidential election results, it is notable that the number of actual voters for most parties and the unaffiliated group sharply decreased. The exception to this trend is the Republican Party, which gained 19,258 more votes than the number of registrations. In contrast, the Democratic Party saw a significant decline, losing 350,372 voters in the election compared to their registrations.*

```{r}
#Prep dataframe for Q1 focusing on Maryland
#Registration Data
md_primary_24 <- read_csv("data/county/county_primary_24.csv")
md_general_24 <- read_csv("data/county/county_general_24.csv")
md_primary_20 <- read_csv("data/county/county_primary_20.csv")
md_primary_20 <- md_primary_20 |>
  rename(County = 1)
md_general_20 <- read_csv("data/county/county_general_20.csv")
#Population Data
state_population <- get_acs(geography="state", 
                       variables = "B01003_001", 
                       year=2022)
md_population <- state_population |>
  slice(21)
#Presidential Election Result
pres_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_1.html"
pres_election_result_html <- pres_election_result_url |>
  read_html() |>
  html_table()
pres_election_result <- pres_election_result_html[[1]]
pres_election_result <- pres_election_result |> clean_names()
pres_election_result <- pres_election_result |>
    slice(-c(34, 35))
```
```{r}
#voters registered for each party in 2024 primary election
total_md_primary_24 <- md_primary_24 |> 
    mutate(OTH = OTH + NLM + LIB) |>
  summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#voters registered for each party in 2024 general election
total_md_general_24 <- md_general_24 |> 
  mutate(OTH = OTH + GRN + NLM + LIB) |> 
summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#voters registered for each party in 2020 primary election
total_md_primary_20 <- md_primary_20 |>
  mutate(OTH = OTH + BAR + WCP) |>
  summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#voters registered for each party in 2020 general election
total_md_general_20 <- md_general_20 |>
  mutate(OTH = OTH + BAR + WCP + GRN + LIB) |>
  summarise(
    DEM_tot = sum(DEM),
    REP_tot = sum(REP),
    OTH_tot = sum(OTH),
    UNA_tot = sum(UNA)
  )
```
```{r}
#How have these numbers changed compared to those in previous elections?*
total_md_primary_24 <- total_md_primary_24 |> 
  mutate(Year = 2024)
total_md_general_24 <- total_md_general_24 |> 
  mutate(Year = 2024)
total_md_primary_20 <- total_md_primary_20 |> 
  mutate(Year = 2020)
total_md_general_20 <- total_md_general_20 |> 
  mutate(Year = 2020)
comparison_general <- bind_rows(total_md_general_24, total_md_general_20) |>
select(Year, everything())
comparison_primary <- bind_rows(total_md_primary_24, total_md_primary_20) |>
select(Year, everything())
```
```{r}
#percentage difference between 2024 and 2022 per each party in geenral elections
comparison_general <- comparison_general|>
    mutate(
    DEM_pct_diff = (DEM_tot[1] - DEM_tot[2]) / DEM_tot[2] * 100,
    REP_pct_diff = (REP_tot[1] - REP_tot[2]) / REP_tot[2] * 100,
    OTH_pct_diff = (OTH_tot[1] - OTH_tot[2]) / OTH_tot[2] * 100,
    UNA_pct_diff = (UNA_tot[1] - UNA_tot[2]) / UNA_tot[2] * 100
  )
```
```{r}
#percentage difference between 2024 and 2022 per each party in primary elections
comparison_primary <- comparison_primary|>
    mutate(
    DEM_pct_diff = (DEM_tot[1] - DEM_tot[2]) / DEM_tot[2] * 100,
    REP_pct_diff = (REP_tot[1] - REP_tot[2]) / REP_tot[2] * 100,
    OTH_pct_diff = (OTH_tot[1] - OTH_tot[2]) / OTH_tot[2] * 100,
    UNA_pct_diff = (UNA_tot[1] - UNA_tot[2]) / UNA_tot[2] * 100
  )
```
```{r}
#voters registered for general 2024 in proportion to population
proportion_md_general_24 <- md_general_24 |>
  summarise(
    total_voters = sum(TOTAL),
    MD_pop = "6161707",
    pct_voters = (total_voters / 6161707) *100
  )
```
```{r}
#voters registered for primary 2024 in proportion to population
proportion_md_primary_24 <- md_primary_24 |>
  summarise(
    total_voters = sum(TOTAL),
    MD_pop = "6161707",
    pct_voters = (total_voters / 6161707) *100
  )
```
```{r}
#*1.3 Compared to the 2024 presidential election result, does the registration pattern match with the result?*
party_elections_results <- pres_election_result |>
  mutate(
    total = as.numeric(gsub(",", "", total))  
  ) |>
  group_by(party) |> 
  summarise(
    Total_votes = sum(total)  
  )
party_elections_results <- party_elections_results |> 
  mutate(
    Total_votes = if_else(party == "Other Candidates", 
                          Total_votes + sum(Total_votes[party %in% c("Green", "Libertarian")]), 
                          Total_votes)
  ) |> 
  filter(!party %in% c("Green", "Libertarian"))
```
```{r}
total_md_general_24_transformed <- total_md_general_24 |> 
  pivot_longer(
    cols = c(DEM_tot, REP_tot, OTH_tot, UNA_tot),  
    names_to = "party",                          
    values_to = "Total_votes"                   
  ) |> 
  mutate(
    party = case_when(                           
      party == "DEM_tot" ~ "Democratic",
      party == "REP_tot" ~ "Republican",
      party == "OTH_tot" ~ "Other Candidates",
      party == "UNA_tot" ~ "Unaffiliated",
      TRUE ~ party
    )
  ) |> 
  select(party, Total_votes)
```
```{r}
comparison_general_elections <- full_join(party_elections_results, total_md_general_24_transformed, by = "party") |>
  rename(Total_election = Total_votes.x, Total_general = Total_votes.y) |>
    mutate(
    Difference = Total_election - Total_general 
    )
```

Q2. At the *Congressional District level*
2.1 How many voters registered for each party in the 2024 primary and general election?
*A: For the 2024 Primary, there are 2,202,830 registrations for Democrats, 994,388 registrations for Republicans, 451,465 registrations for Unaffiliated, and 31,418 registrations for other parties. For 2024 General, there are 2,211,316 registrations for Democrats, 1,007,384 registrations for Republicans, 907,638 registrations for Unaffiliated, and 78,234 registrations for other parties.*
2.1.1 How have these numbers changed compared to those in previous elections?
*A: For the 2024 and 2020 Primary, the registration for Democrats decreased by 1.24%, Republicans decreased by 1.45%, Unaffiliated increased by 40.24%, and other parties increased by 44.96%. For 2024 and 2020 General, the registration for Democrats decreased by 2.28%, Republicans decreased by 0.83%, Unaffiliated increased by 18.37%, and other parties increased by 21.55%.*
2.2 Are there any patterns for voter registration from every district?
*A: For both Primary and General 2024, in most districts, there are registrations for Democrats more than registration for Republicans.*
2.2.1 Are there any districts that don't fall into this pattern?
*A: There is only District 1 where registration for Democrats is less than that of Republicans.*
2.2.2 How are those districts different from others in terms of location, demography, economy, etc.?
*A:.....* 
2.3 Compared to the total population of each district, what is the proportion of voters who register for the election? 
*A: For both Primary and General, District 5 has the most proportion of registered voters which are 69.26% for Primary and 74.12% for General. In the other hand, District 4 has the least proportion of registered voters which are 52.02% for Primary and 58.34% for General* 
2.3.1 What are the differences between district having large and small proportions of voter registration?
*A:.....* 
2.4 Compared to the representative race result, does the registration pattern match with the result?
*A: Yes, the General registration pattern in every district is match with the pattern of the representative race result. There are more percentage of people who registered and votes for Democrats than those for Republicans in District 2 to 8, and on the contrary, there are more percentage of people who registered and votes for Republicans than those for Democrats in District 1.* 

```{r}
#Prep dataframe for Q2 focusing on congressional district level

#Registration Data
congress_primary_24 <- read_csv("data/congress_level/congress_primary_24.csv")
congress_general_24 <- read_csv("data/congress_level/congress_general_24.csv")
congress_primary_20 <- read_csv("data/congress_level/congress_primary_20.csv")
congress_primary_20 <- congress_primary_20 |>
  rename(District = 1) |>
  rename(county = 2) |>
  rename(BAR = 3) |>
  rename(DEM = 4) |>
  rename(OTH = 5) |>
  rename(REP = 6) |>
  rename(UNA = 7) |>
  rename(WCP = 8) |>
  rename(total = 9) 
congress_general_20 <- read_csv("data/congress_level/congress_general_20.csv")

#Population Data
congress_population <- get_acs(geography="congressional district", 
                       variables = "B01003_001", state = "MD", 
                       year=2022)

#Representative Election Result
representative_election_result_url <- "https://elections.maryland.gov/elections/2024/general_Results/gen_results_2024_3.html"
representative_election_result_html <- representative_election_result_url |>
  read_html() |>
  html_table()
representative_election_result1 <- representative_election_result_html[[1]]
representative_election_result2 <- representative_election_result_html[[2]]
representative_election_result3 <- representative_election_result_html[[3]]
representative_election_result4 <- representative_election_result_html[[4]]
representative_election_result5 <- representative_election_result_html[[5]]
representative_election_result6 <- representative_election_result_html[[6]]
representative_election_result7 <- representative_election_result_html[[7]]
representative_election_result8 <- representative_election_result_html[[8]]
```
```{r}
#Primary
#2024
count_congress_primary_24 <- congress_primary_24 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    LIB = sum(LIB),
    OTH = sum(OTH),
    UNA = sum(UNA),
    NLM = sum(NLM)
  ) |>
  mutate(Other = LIB+OTH+NLM) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM24 = 1, REP24 =2, UNA24 = 3, Other24 = 4)
#2020
count_congress_primary_20 <- congress_primary_20 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    OTH = sum(OTH),
    UNA = sum(UNA),
    BAR = sum(BAR),
    WCP = sum(WCP)
  ) |>
  mutate(Other = OTH+BAR+WCP) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM20 = 1, REP20 =2, UNA20 = 3, Other20 = 4)
#Find the change
compare_congress_primary_24_20 <- bind_cols(count_congress_primary_24, count_congress_primary_20) |>
  mutate(DEM_change = ((DEM24-DEM20)/DEM20)*100) |>
  mutate(REP_change = ((REP24-REP20)/REP20)*100) |>
  mutate(UNA_change = ((UNA24-UNA20)/UNA20)*100) |>
  mutate(Other_change = ((Other24-Other20)/Other20)*100)

#General
#2024
count_congress_general_24 <- congress_general_24 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    LIB = sum(LIB),
    OTH = sum(OTH),
    UNA = sum(UNA),
    NLM = sum(NLM),
    GRN = sum(GRN)
  ) |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM24 = 1, REP24 =2, UNA24 = 3, Other24 = 4)
#2020
count_congress_general_20 <- congress_general_20 |>
  summarise(
    DEM = sum(DEM),
    REP = sum(REP),
    BAR = sum(BAR),
    GRN = sum(GRN),
    LIB = sum(LIB),
    WCP = sum(WCP),
    OTH = sum(OTH),
    UNA = sum(UNA)
  ) |>
  mutate(Other = OTH+BAR+GRN+LIB+WCP) |>
  select(DEM, REP, UNA, Other) |>
  rename(DEM20 = 1, REP20 =2, UNA20 = 3, Other20 = 4)
#Find the change
compare_congress_general_24_20 <- bind_cols(count_congress_general_24, count_congress_general_20) |>
  mutate(DEM_change = ((DEM24-DEM20)/DEM20)*100) |>
  mutate(REP_change = ((REP24-REP20)/REP20)*100) |>
  mutate(UNA_change = ((UNA24-UNA20)/UNA20)*100) |>
  mutate(Other_change = ((Other24-Other20)/Other20)*100)

#Find Pattern
#Primary24
pattern_congress_primary_24 <- congress_primary_24 |>
  mutate(Other = LIB+OTH+NLM) |>
  select(District, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100)
#General24
pattern_congress_general_24 <- congress_general_24 |>
  mutate(Other = LIB+OTH+NLM+GRN) |>
  select(District, DEM, REP, UNA, Other, TOTAL) |>
  mutate(DEM_per = (DEM/TOTAL)*100) |>
  mutate(REP_per = (REP/TOTAL)*100) |>
  mutate(UNA_per = (UNA/TOTAL)*100) |>
  mutate(Other_per = (Other/TOTAL)*100)

#Compare to Population
#Primary24
total_congress_primary_24 <- congress_primary_24 |>
  select(District, TOTAL)
cleaned_congress_population <- congress_population |>
  rename(population = 4) |>
  select(4)
regis_population_congress_primary_24 <- bind_cols(total_congress_primary_24, cleaned_congress_population) |>
  mutate(Percent = (TOTAL/population)*100)
#General24
total_congress_general_24 <- congress_general_24 |>
  select(District, TOTAL)
regis_population_congress_general_24 <- bind_cols(total_congress_general_24, cleaned_congress_population) |>
  mutate(Percent = (TOTAL/population)*100)

#Compare to the result
#Clean the result
representative_election_result1 <- representative_election_result1 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result2 <- representative_election_result2 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result3 <- representative_election_result3 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result4 <- representative_election_result4 |>
  slice(-4) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, Other_percent_result = 4) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(Other_percent_result = as.numeric(gsub("%","", Other_percent_result)))
representative_election_result5 <- representative_election_result5 |>
  slice(-4) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, Other_percent_result = 4) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(Other_percent_result = as.numeric(gsub("%","", Other_percent_result)))
representative_election_result6 <- representative_election_result6 |>
  slice(-4) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, Other_percent_result = 4) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(Other_percent_result = as.numeric(gsub("%","", Other_percent_result)))
representative_election_result7 <- representative_election_result7 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result8 <- representative_election_result8 |>
  slice(-5) |>
  select(1, 8) |>
  pivot_longer(cols = -Name, names_to = "Party", values_to = "values") |>
  pivot_wider(names_from = Name, values_from = values) |>
  rename(DEM_percent_result = 2, REP_percent_result = 3, x = 4, y = 5) |>
  mutate(DEM_percent_result = as.numeric(gsub("%","", DEM_percent_result)))|>
  mutate(REP_percent_result = as.numeric(gsub("%","", REP_percent_result)))|>
  mutate(x = as.numeric(gsub("%","", x))) |>
  mutate(y = as.numeric(gsub("%","", y))) |>
  mutate(Other_percent_result = x+y) |>
  select(1,2,3,6)
representative_election_result <- bind_rows(list(representative_election_result1, representative_election_result2, representative_election_result3, representative_election_result4, representative_election_result5, representative_election_result6, representative_election_result7, representative_election_result8)) |>
  select(2,3,4)
#Merge with the registration
congress_general_regis_result <- bind_cols(pattern_congress_general_24, representative_election_result) |>
  select(-2,-3,-4,-5,-6)
```


































































