---
title: "lab_05"
author: "derek willis"
date: "2024-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# load the tidyverse
library(tidyverse)
```

## Getting Started

This week we'll be combining and joining data to make it more useful and to ask some questions about it. This week, we'll revisit our new Maryland voters data and try to join it with contributions to the Harris campaign from Maryland donors made via ActBlue. Can we find new voters who are not registered Democrats who gave to Harris after she became a presidential candidate on July 21? What, if anything, can we say about those people?

We'll be doing something similar to what Ence Morse did for [this Washington Post story](https://www.washingtonpost.com/politics/2024/09/24/kamala-harris-younger-donors/?pwapi_token=eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJyZWFzb24iOiJnaWZ0IiwibmJmIjoxNzI3MzIzMjAwLCJpc3MiOiJzdWJzY3JpcHRpb25zIiwiZXhwIjoxNzI4NzA1NTk5LCJpYXQiOjE3MjczMjMyMDAsImp0aSI6ImIwYWMxNTk4LWMwMDgtNDE5Ny04MjAyLWVmMWNjZWFmYWZiMiIsInVybCI6Imh0dHBzOi8vd3d3Lndhc2hpbmd0b25wb3N0LmNvbS9wb2xpdGljcy8yMDI0LzA5LzI0L2thbWFsYS1oYXJyaXMteW91bmdlci1kb25vcnMvIn0.3-s0tBkMV7Y9DD21CBsKMlaliGlOKT2egCO-UFT735Y), just on a smaller scale.

First, load two CSV contribution files in the `data` folder, one each for donations made in July and August, and then combine them into a single dataframe.

```{r}
# load the Harris contribution data files and combine them
august_harris <- read_csv("data/august_harris.csv")
july_harris <- read_csv("data/july_harris.csv")
july_august_combined = bind_rows(list(august_harris, july_harris))

```

Next, let's read in the new voters (people who registered in 2024):

```{r}
# load the new voters data
new_voters2024 <- read_csv("data/new_voters.csv")
```

Next join the Harris donations to the voter data. Remember, people's names are common, __so you'll want to use name columns and address information__. You'll need to figure out which columns to use from both dataframes. The new_voters data has both a single Address column and various parts of the address, such as HouseNumber. Make a plan for picking your join columns - try out some combinations - and save the results to a new dataframe.

```{r}
# join the combined Harris file to the new voters
# you will need to determine what columns to use
new_data <- july_august_combined |> inner_join(new_voters2024, join_by(address_one==Address, last_name==LastName, first_name==FirstName))

new_data_address <- july_august_combined |> inner_join(new_voters2024, join_by(address_one==Address))

new_data_lastfirst <- july_august_combined |> inner_join(new_voters2024, join_by(last_name==LastName, first_name==FirstName))
```

## Answer questions

Q1. Explain why you used the columns you did to make the join, and why that result is better than other ways you tried. Are there any problems with the join you made, such as the possibility of false positives?

A1. The combination of address, last name, and first name is most likely to uniquely identify an individual than any of those fields alone. It only returns 365 observations. A possoble false positive here could be multiple people with the same name living in a large communal residence like a dorm. Another one could be parent-child names differing only by Sr. and Jr or having a different middle name, living in the same house. The second join (new_data_address) uses only the address to produce the join and there are 2223 observations using that. It's more prone to false positives, as different people can live at the same address. Using last and first name returns 8251 observations. People are very likely to have the same full name. 

Q2. Based on your join, how many donations to Harris have come from people who registered this year? How many unique people does that represent, and which donors have given the largest _number_ of donations since July 20? 

A2. 365 donations in total. There are 270 unique new registered donors. 	
SOPHIA HOFFER-PERKINS, BRITTANY MUSOLINO, ERIC SNIPE each have 7 donations, making them the donors with the largest number of donations since July 20. The data only has people with donations after july 20. 


```{r}
unique_donors <- new_data |> group_by(last_name, first_name)|>
 summarise(count = n())|>
  arrange(desc(count))
 # summarise(n_distinct(first_name, last_name))
#glimpse
```

Q3. What are the party and gender breakdowns of those donors? How many donations has Harris gotten from non-Democrats?

A3.  Sophia Hoffer Perkins and Brittany Musolino are female. Eric Snipe is male. They're all democrats. 43 donations are from non democrats

```{r}
new_data |> 
  group_by(last_name, first_name, address_one) |> 
  summarise(party = Party,
            gender = Gender,
            count = n()) |> 
  arrange(desc(count))

nondemocrat_donations <- new_data |>
  filter(Party != "DEM") |>
  nrow()
```


Q4. Choose your own adventure: I want you to come up with a question you could ask of one of the dataframes we've created for this lab. It could involve just the new voters, or the donors or the joined dataframe, but it cannot be a variation of a previous question (so, for example, you can't ask, "How many Republicans have given to Harris"). First write the question and then try to answer it by writing code. If you're not sure of the exact code, try googling (or ask me for advice) but, at a minimum, describe what you think you'd have to do to answer it and what the results would look like. That can help.

A4: I am curious to know the kind of jobs people have who are donating to Harris; campaign. Q: What is the most common occupation among both new voters and july/august voters combined who have donated to Harris?"
```{r}
occupation_analysis_nv <- new_data |> #new voters' occupations
  group_by(occupation)|>
  summarise(count = n())|>
  arrange(desc(count))

occupation_analysis_all <- july_august_combined |> #july august voters combined occupations
  group_by(occupation)|>
  summarise(count = n())|>
  arrange(desc(count))

```
